---
title: "Course Project: Shiny Application and Reproducible Pitch"
author: "Nidhi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Diamond Analysis Report

Diamond Data Analysis Report enabled developing data products using shiny apps being embeded within RMarkDown.

Using PageUp and PageDown keys all the slides can be traversed.

## Application Overview

- Diamonds Properties can be analysed using several Parameters.
- Carat,Price,sample size,color Depth are some of the common parameters.
- This Report enables to pick the right parameters for the best Diamond selection.


## Slide with Complete Reporting Appliaction

Click the Below Link for the Application.

https://nidhi31.shinyapps.io/Week4/?_ga=2.196753004.195944286.1595842430-766815401.1593517691

Click the link below for compiled project files on gitub repo.

https://github.com/snidhi92/datasciencecoursera/tree/master/Developing%20Data%20Products/Week4

## Data Used
The data used for this application is diamonds data set, which is part of ggplot2 package. Containing information about 53940 diamonds with 10 variables.

```{r, echo=TRUE}
library("ggplot2")
head(diamonds)
```